The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey

### 2020-05-16 DES_RET_Hosts_submit.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib,]{mnras}
\usepackage{lineno}
\linenumbers
% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{booktabs}   % Make tables directly in pandas
\usepackage{longtable}  % Add longer tables
\usepackage{rotating}
\usepackage{appendix}
\usepackage{float}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{wrapfig}
\usepackage{ctable}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\captionsetup{compatibility=false}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[RET host galaxies in DES]{The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[DES Collaboration]{
\parbox{\textwidth}{
\Large
P.~Wiseman,$^{1}$
M.~Pursiainen,$^{1}$
M.~Childress,$^{1}$
E.~Swann,$^{2}$
M.~Smith,$^{1}$
L.~Galbany,$^{3}$
C.~Lidman,$^{4}$
T.~M.~Davis,$^{5}$
C.~P.~Guti\'errez,$^{1}$
A.~M\"oller,$^{6}$
B.~P.~Thomas,$^{2}$
C.~Frohmaier,$^{2}$
R.~J.~Foley,$^{7}$
S.~R.~Hinton,$^{5}$
L.~Kelsey,$^{1}$
R.~Kessler,$^{8,9}$
G.~F.~Lewis,$^{10}$
M.~Sako,$^{11}$
D.~Scolnic,$^{12}$
M.~Sullivan,$^{1}$
M.~Vincenzi,$^{2}$
T.~M.~C.~Abbott,$^{13}$
M.~Aguena,$^{14,15}$
S.~Allam,$^{16}$
J.~Annis,$^{16}$
E.~Bertin,$^{17,18}$
S.~Bhargava,$^{19}$
D.~Brooks,$^{20}$
D.~L.~Burke,$^{21,22}$
A.~Carnero~Rosell,$^{23}$
D.~Carollo,$^{24}$
M.~Carrasco~Kind,$^{25,26}$
J.~Carretero,$^{27}$
M.~Costanzi,$^{28,29}$
L.~N.~da Costa,$^{15,30}$
H.~T.~Diehl,$^{16}$
P.~Doel,$^{20}$
S.~Everett,$^{7}$
J.~Frieman,$^{16,9}$
J.~Garc\'ia-Bellido,$^{31}$
E.~Gaztanaga,$^{32,33}$
K.~Glazebrook,$^{34}$
D.~Gruen,$^{35,21,22}$
R.~A.~Gruendl,$^{25,26}$
J.~Gschwend,$^{15,30}$
G.~Gutierrez,$^{16}$
K.~Honscheid,$^{36,37}$
D.~J.~James,$^{38}$
K.~Kuehn,$^{39,40}$
N.~Kuropatkin,$^{16}$
M.~Lima,$^{14,15}$
M.~A.~G.~Maia,$^{15,30}$
J.~L.~Marshall,$^{41}$
P.~Martini,$^{36,42}$
F.~Menanteau,$^{25,26}$
R.~Miquel,$^{43,27}$
A.~Palmese,$^{16,9}$
F.~Paz-Chinch\'{o}n,$^{44,26}$
A.~A.~Plazas,$^{45}$
A.~K.~Romer,$^{19}$
E.~Sanchez,$^{23}$
M.~Schubnell,$^{46}$
S.~Serrano,$^{32,33}$
I.~Sevilla-Noarbe,$^{23}$
N.~E.~Sommer,$^{4}$
E.~Suchyta,$^{47}$
M.~E.~C.~Swanson,$^{26}$
G.~Tarle,$^{46}$
B.~E.~Tucker,$^{4}$
D.~L.~Tucker,$^{16}$
T.~N.~Varga,$^{48,49}$
and A.~R.~Walker$^{13}$
\begin{center} (DES Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{Affiliations are listed at the end of the paper}
}
% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a distinct class of astrophysical event. They are characterised by lightcurves that decline much faster than common classes supernovae (SNe), span vast ranges in peak luminosity and can be seen to redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the first substantial analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs, indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

In the standard paradigm of stellar evolution, stars with a zero-age main sequence (ZAMS) mass above $8M_{\sun}$ are believed to explode as a result of a catastrophic collapse of their iron cores and are known as core-collapse supernovae (CCSNe). CCSNe can be split into observationally-determinded subclasses based on their lightcurve and spectral evolution: SNe II display hydrogen features in their spectra \citep{Minkowski1941}, and are thought to occur in stars that retain a large fraction of their hydrogen envelope. Conversely, SNe Ib and Ic do not show signatures of hydrogen \citep[e.g.][]{Filippenko1997} and are thus referred to collectively as stripped-envelope SNe (SESNe), having undergone a partial removal of their outer atmospheres. The SN IIb subclass, which shows hydrogen only at early epochs that disappears after a few weeks \citep{Filippenko1988}, is also commonly grouped along with SESNe. \replyrick{SNe IIn display much narrower hydrogen emission lines when compared to standard SNe II \citep{Schlegel1990}. The narrow emission originates from the ejecta impacting on slow-moving circumstellar material (CSM)}. Since the turn of the century, observations of CCSNe, whose lightcurves are primarily powered by the radioactive decay of freshly synthesised Ni-56, have been supplemented by rarer, more exotic transient classes.

 Long duration gamma-ray bursts (LGRBs), although first discovered in the 1960s \citep{Klebesadel1973}, were unequivocally linked to collapsing massive stars through their associations with broad-lined type Ic SNe \citep{Galama1998,Hjorth2003}. Thought to be caused by accretion onto a newly-formed black hole at the centre of a collapsing, rapidly-rotating massive star \citep[e.g.][]{Woosley1993,Woosley2006a,Woosley2006b}, LGRBs comprise roughly $1\%$ of all SNe Ic, which themselves make up only 15\% of all CCSNe \citep{Kelly2012,Graham2016}. \replyrick{Another} exotic class of SNe is the particularly bright superluminous supernovae (SLSNe; e.g. \citealt{Quimby2011, Gal-Yam2012}). Originally grouped due to their slowly-evolving lightcurves and extreme luminosity (peaking at $M_B < -21$~mag; 10-100 times brighter than \replyrick{typical} CCSNe), recent observations have revealed a continuum of spectroscopically similar objects with peaks as faint as $M_B \sim -19$~mag \citep{DeCia2018,Lunnan2018,Angus2019}, \replyrick{which overlaps with} the bright end of the CCSN luminosity function \citep{Li2011}. The lightcurve evolution of SLSNe is not well described by models of Ni-56 decay, with the most popular alternative hypothesis being the magnetic coupling of the ejecta with the spin down of a newly formed, rapidly rotating magnetar.

 Along with observations of the transients themselves, \replyrick{studies of} host galaxies are frequently \replyrick{used to make strong inferences about the progenitor stars and explosion mechanisms}. CCSNe are confined almost exclusively to galaxies hosting recent or ongoing star formation, due to their origin from massive stars. There are correlations between the expected progenitor mass of different sub-classes of CCSNe and host galaxy properties. On average, SESNe reside in galaxies with higher specific star-formation rates (sSFRs; \citealt{James2006,Kelly2008}), while studies of the local environments tend to show that SESNe explode closer to \HII~regions than SNe II, indicating that the progenitors are younger and more massive than the various sub-classes of hydrogen-rich SNe II  \citep[e.g.][]{Anderson2012,Galbany2018}. More extreme events tend to occur in galaxies low in mass and high in sSFR, with both GRBs \citep[e.g.][]{Fruchter2006,LeFloch2006,Levesque2010,Kruehler2015,Vergani2015,Perley2016b,Palmerio2019,Taggart2019} and to an even greater degree SLSNe \citep[e.g.][]{Neill2011,Lunnan2014,Leloudas2015,Angus2016,Schulze2018,Taggart2019} exhibiting this association.

 The chemical composition of the interstellar medium (ISM) is an important consideration when comparing host galaxy properties. While it does not appear to play a significant role in the relative production of CCSNe (although there are some trends, with SESNe typically found in slightly less metal-rich environments than SNe II; \citealt{Galbany2018}), it appears to be vitally important in the production of LGRBs and SLSNe. Theory predicts that the production of a LGRB should only be possible in stars with a metallicity of $Z/Z_{\sun}\leq 0.3$ \citep{Woosley1993}  in order for the likely Wolf-Rayet or blue supergiant progenitors not to lose their outer atmospheres through metal-driven winds, thus conserving sufficient angular momentum to power the black-hole-driven jet or rapidly rotating magnetar. Many LGRB host galaxy studies have indeed revealed a metallicity threshold between 0.5 and 1 times the solar value \citep[e.g.][]{Stanek2006,Modjaz2008,Kruehler2015,Perley2016b,Japelj2016,Vergani2017}, although this is not a rigid \replyrick{threshold}.  
SLSN host galaxies also appear to be lower in metallicity than would be expected for their stellar mass, with a suppression of SLSN production \replyrick{above a value of about} half-solar \citep{Lunnan2014,Chen2016a,Perley2016c}. \replyrick{Like LGRBs, SLSNe} also require a particularly high sSFR, suggesting that they are explosions of very young, rapidly rotating massive stars.

Recently, inspection of high-cadence, large-area survey data sets have revealed more exotic transients that are \replyrick{difficult} to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly like the PS1 sample. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. RETs typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by expanding blackbodies, although a handful show declining photospheric radii from the \replyrick{first detection}. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and \replyrick{from X-rays through to radio wavelengths} did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}) . There are many diverse explanations for the power source of AT2018cow touted in the literature, including: magnetars \citep{Mohan2020}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Uno2020}.
Other nearby rapid transients include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the \replyrick{accretion-induced collapse of a white dwarf or a white dwarf-neutron star merger}, and KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material.  It is currently unclear whether these transients represent the local analogues of the DES and PS\replyrick{1} RETs.

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} by adding the 5th and final \replyrick{season} of DES-SN observations as well as more refined \replyrick{selection} techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties \replyrick{and} compare them to \replyrick{samples of host galaxies of} CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

@@

\replyrick{The Dark Energy Survey (DES) makes use of the Dark Energy Camera (DECam; \citealt{Flaugher2015}) to survey $5000\deg^2$ of the southern sky between 2013 and 2019. The supernova programme (DES-SN), designed primarily to detect type Ia SNe (SNe Ia) for cosmological measurements \citep{Bernstein2012}, consisted of five 6-month seasons of approximately 7 day cadence \citep{DAndrea2018}} in ten single pointings, known as the SN fields which cover a total area of $\sim27 \deg^2$. Of these fields, two were observed with longer exposure times and are referred to as deep fields; the remaining eight are labelled shallow. Transient detection was performed by difference imaging using a custom pipeline \citep{Kessler2015}. Transient candidates were vetted via machine-learning techniques \citep{Goldstein2015}, leading to $\sim 30,000$ viable supernova candidates over the full five seasons. A full description of the search for rapid transients in DES can be found in \citetalias{Pursiainen2018}.

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to \replyrick{quantify the completeness of the sample}. For instance, due to the large redshift range within the sample it is entirely possible that distant events \replyrick{with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events and low redshift may evolve on timescales quicker than the survey cadence}. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, \replyrick{we perform the search} in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution (rise to peak in $\lesssim15$ days) and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient, they still effectively distinguish the the fast events from traditional SN types. We thus attempt to select a sample based on observed rise times and colours at peak brightness.

\subsubsection{Gaussian processed lightcurves}
Using observed photometric data points directly to infer rise time and colour has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights in the `deep' fields, thus making it impossible to measure the peak colour accurately. Measuring rise times of 10-15 days is difficult to perform with a one week cadence when it has to be done without fitting a lightcurve model. To negate this issue, we use interpolate the lightcurves of all DES-SN candidates using Gaussian Processes (GP) as presented in \citet{Pursiainen2020}. The interpolated lightcurves have a 0.5 day cadence and every epoch and band has a flux value and an associated uncertainty.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib,]{mnras}
\usepackage{lineno}
%\linenumbers
% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{booktabs}   % Make tables directly in pandas
\usepackage{longtable}  % Add longer tables
\usepackage{rotating}
\usepackage{appendix}
\usepackage{float}
\usepackage[dvipsnames]{xcolor}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{multicol}
%\usepackage{wrapfig}
\usepackage{ctable}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\captionsetup{compatibility=false}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[RET host galaxies in DES]{The Host Galaxies of Rapidly Evolving Transients in the Dark Energy Survey}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[P. S. Wiseman]{
\parbox{\textwidth}{
\Large
P.~Wiseman,$^{1}$
M.~Pursiainen,$^{1}$
M.~Childress,$^{1}$
E.~Swann,$^{2}$
M.~Smith,$^{1}$
L.~Galbany,$^{3}$
C.~Lidman,$^{4}$
T.~M.~Davis,$^{5}$
C.~P.~Guti\'errez,$^{1}$
A.~M\"oller,$^{6}$
B.~P.~Thomas,$^{2}$
C.~Frohmaier,$^{2}$
R.~J.~Foley,$^{7}$
S.~R.~Hinton,$^{5}$
L.~Kelsey,$^{1}$
R.~Kessler,$^{8,9}$
G.~F.~Lewis,$^{10}$
M.~Sako,$^{11}$
D.~Scolnic,$^{12}$
M.~Sullivan,$^{1}$
M.~Vincenzi,$^{2}$
T.~M.~C.~Abbott,$^{13}$
M.~Aguena,$^{14,15}$
S.~Allam,$^{16}$
J.~Annis,$^{16}$
E.~Bertin,$^{17,18}$
S.~Bhargava,$^{19}$
D.~Brooks,$^{20}$
D.~L.~Burke,$^{21,22}$
A.~Carnero~Rosell,$^{23}$
D.~Carollo,$^{24}$
M.~Carrasco~Kind,$^{25,26}$
J.~Carretero,$^{27}$
M.~Costanzi,$^{28,29}$
L.~N.~da Costa,$^{15,30}$
H.~T.~Diehl,$^{16}$
P.~Doel,$^{20}$
S.~Everett,$^{7}$
P.~Fosalba,$^{31,32}$
J.~Frieman,$^{16,9}$
J.~Garc\'ia-Bellido,$^{33}$
E.~Gaztanaga,$^{31,32}$
K.~Glazebrook,$^{34}$
D.~Gruen,$^{35,21,22}$
R.~A.~Gruendl,$^{25,26}$
J.~Gschwend,$^{15,30}$
G.~Gutierrez,$^{16}$
D.~L.~Hollowood,$^{7}$
K.~Honscheid,$^{36,37}$
D.~J.~James,$^{38}$
K.~Kuehn,$^{39,40}$
N.~Kuropatkin,$^{16}$
M.~Lima,$^{14,15}$
M.~A.~G.~Maia,$^{15,30}$
J.~L.~Marshall,$^{41}$
P.~Martini,$^{36,42}$
F.~Menanteau,$^{25,26}$
R.~Miquel,$^{43,27}$
A.~Palmese,$^{16,9}$
F.~Paz-Chinch\'{o}n,$^{44,26}$
A.~A.~Plazas,$^{45}$
A.~K.~Romer,$^{19}$
E.~Sanchez,$^{23}$
V.~Scarpine,$^{16}$
M.~Schubnell,$^{46}$
S.~Serrano,$^{31,32}$
I.~Sevilla-Noarbe,$^{23}$
N.~E.~Sommer,$^{4}$
E.~Suchyta,$^{47}$
M.~E.~C.~Swanson,$^{26}$
G.~Tarle,$^{46}$
B.~E.~Tucker,$^{4}$
D.~L.~Tucker,$^{16}$
T.~N.~Varga,$^{48,49}$
and A.~R.~Walker$^{13}$
\begin{center} (DES Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{Affiliations are listed at the end of the paper}
}
% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a distinct class of astrophysical event. They are characterised by lightcurves that decline much faster than common classes supernovae (SNe), span vast ranges in peak luminosity and can be seen to redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the most comprehensive analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs, indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

In the standard paradigm of stellar evolution, stars with a zero-age main sequence (ZAMS) mass above $8M_{\sun}$ are believed to explode as a result of a catastrophic collapse of their iron cores and are known as core-collapse supernovae (CCSNe). CCSNe can be split into observationally-determined subclasses based on their lightcurve and spectral evolution: SNe II display hydrogen features in their spectra \citep{Minkowski1941}, and are thought to occur in stars that retain a large fraction of their hydrogen envelope. Conversely, SNe Ib and Ic do not show signatures of hydrogen \citep[e.g.][]{Filippenko1997} and are thus referred to collectively as stripped-envelope SNe (SESNe), having undergone a partial removal of their outer atmospheres. The SN IIb subclass, which shows hydrogen only at early epochs that disappears after a few weeks \citep{Filippenko1988}, is also commonly grouped along with SESNe. SNe IIn display much narrower hydrogen emission lines when compared to standard SNe II \citep{Schlegel1990}. The narrow emission originates from the ejecta impacting on slow-moving circumstellar material (CSM). Since the turn of the century, observations of CCSNe, whose lightcurves are primarily powered by the radioactive decay of freshly synthesised Ni-56, have been supplemented by rarer, more exotic transient classes.

 Long duration gamma-ray bursts (LGRBs), although first discovered in the 1960s \citep{Klebesadel1973}, were unequivocally linked to collapsing massive stars through their associations with broad-lined type Ic SNe \citep{Galama1998,Hjorth2003}. Thought to be caused by accretion onto a newly-formed black hole at the centre of a collapsing, rapidly-rotating massive star \citep[e.g.][]{Woosley1993,Woosley2006a,Woosley2006b}, LGRBs comprise roughly $1\%$ of all SNe Ic, which themselves make up only 15\% of all CCSNe \citep{Kelly2012,Graham2016}. Another exotic class of SNe is the particularly bright superluminous supernovae (SLSNe; e.g. \citealt{Quimby2011, Gal-Yam2012}). Originally grouped due to their slowly-evolving lightcurves and extreme luminosity (peaking at $M_B < -21$~mag; 10-100 times brighter than typical CCSNe), recent observations have revealed a continuum of spectroscopically similar objects with peaks as faint as $M_B \sim -19$~mag \citep{DeCia2018,Lunnan2018,Angus2019}, which overlaps with the bright end of the CCSN luminosity function \citep{Li2011}. The lightcurve evolution of SLSNe is not well described by models of Ni-56 decay, with the most popular alternative hypothesis being the magnetic coupling of the ejecta with the spin down of a newly formed, rapidly rotating magnetar.

 Along with observations of the transients themselves, studies of host galaxies are frequently used to make strong inferences about the progenitor stars and explosion mechanisms. CCSNe are confined almost exclusively to galaxies hosting recent or ongoing star formation, due to their origin from massive stars. There are correlations between the expected progenitor mass of different sub-classes of CCSNe and host galaxy properties. On average, SESNe reside in galaxies with higher specific star-formation rates (sSFRs; \citealt{James2006,Kelly2008}), while studies of the local environments tend to show that SESNe explode closer to \HII~regions than SNe II, indicating that the progenitors are younger and more massive than the various sub-classes of hydrogen-rich SNe II  \citep[e.g.][]{Anderson2012,Galbany2018}. More extreme events tend to occur in galaxies low in mass and high in sSFR, with both LGRBs \citep[e.g.][]{Fruchter2006,LeFloch2006,Levesque2010,Kruehler2015,Vergani2015,Perley2016b,Palmerio2019,Taggart2019} and to an even greater degree SLSNe \citep[e.g.][]{Neill2011,Lunnan2014,Leloudas2015,Angus2016,Schulze2018,Taggart2019} exhibiting this association.

 The chemical composition of the interstellar medium (ISM) is an important consideration when comparing host galaxy properties. While it does not appear to play a significant role in the relative production of CCSNe (although there are some trends, with SESNe typically found in slightly less metal-rich environments than SNe II; \citealt{Galbany2018}), it appears to be vitally important in the production of LGRBs and SLSNe. Theory predicts that the production of a LGRB should only be possible in stars with a metallicity of $Z/Z_{\sun}\leq 0.3$ \citep{Woosley1993}  in order for the likely Wolf-Rayet or blue supergiant progenitors not to lose their outer atmospheres through metal-driven winds, thus conserving sufficient angular momentum to power the black-hole-driven jet or rapidly rotating magnetar. Many LGRB host galaxy studies have indeed revealed a metallicity threshold between 0.5 and 1 times the solar value \citep[e.g.][]{Stanek2006,Modjaz2008,Kruehler2015,Perley2016b,Japelj2016,Vergani2017}, although this is not a rigid threshold.  
SLSN host galaxies also appear to be lower in metallicity than would be expected for their stellar mass, with a suppression of SLSN production above a value of about half-solar \citep{Lunnan2014,Chen2016a,Perley2016c}. Like LGRBs, SLSNe also require a particularly high sSFR, suggesting that they are explosions of very young, rapidly rotating massive stars.

Recently, inspection of high-cadence, large-area survey data sets have revealed more exotic transients that are difficult to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly like the PS1 sample. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. RETs typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by expanding blackbodies, although a handful show declining photospheric radii from the first detection. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and from X-rays through to radio wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}) . There are many diverse explanations for the power source of AT2018cow touted in the literature, including: magnetars \citep{Mohan2020}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Uno2020}.
Other nearby rapid transients include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the accretion-induced collapse of a white dwarf or a white dwarf-neutron star merger, and KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material.  It is currently unclear whether these transients represent the local analogues of the DES and PS1 RETs.

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} by adding the 5th and final season of DES-SN observations as well as more refined selection techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties and compare them to samples of host galaxies of CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

@@

The Dark Energy Survey (DES) makes use of the Dark Energy Camera (DECam; \citealt{Flaugher2015}) to survey $5000\deg^2$ of the southern sky between 2013 and 2019. The supernova programme (DES-SN), designed primarily to detect type Ia SNe (SNe Ia) for cosmological measurements \citep{Bernstein2012}, consisted of five 6-month seasons of approximately 7 day cadence in ten single pointings, known as the SN fields which cover a total area of $\sim27 \deg^2$. Of these fields, two were observed with longer exposure times and are referred to as deep fields with remaining eight known as shallow fields; the resulting approximate single-visit depths are 23.5 mag (shallow) and 24.5 mag (deep) in the DES $r$-band. Transient detection was performed by difference imaging using a custom pipeline \citep{Kessler2015}. Transient candidates were vetted via machine-learning techniques \citep{Goldstein2015}, leading to $\sim 30,000$ viable supernova candidates over the full five seasons. DES-SN included an extensive spectroscopic follow-up program to identify transients and measure spectroscopic redshifts of host galaxies \citep{DAndrea2018}. A full description of the search for rapid transients in DES can be found in \citetalias{Pursiainen2018}.

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to quantify the completeness of the sample. For instance, due to the large redshift range within the sample it is entirely possible that distant events with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events and low redshift may evolve on timescales quicker than the survey cadence. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, we perform the search in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution (rise to peak in $\lesssim15$ days) and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient, they still effectively distinguish the the fast events from traditional SN types. We thus attempt to select a sample based on observed rise times and colours at peak brightness.

\subsubsection{Gaussian processed lightcurves}
Using observed photometric data points directly to infer rise time and colour has several problems that can be improved. For one, measuring peak colour is problematic: DES-SN did not always observe $g$- and $r$-bands on the same or even consecutive nights in the `deep' fields, thus adding larger uncertainty in the peak colour estimate. We do not have a light curve model for RETs, and therefore  measuring a 10-15 day rise time is difficult with a one week  cadence. Rather than fitting with a physically motivated model, we instead interpolate the lightcurves of all DES-SN candidates using Gaussian Processes (GP) as presented in \citet{Pursiainen2020}. The interpolated lightcurves have a 0.5 day cadence and every epoch and band has a flux value and an associated uncertainty.

\subsubsection{Photometric definition of RETs \label{subsubsec:trise_gr}}
To make an improved selection of RETs we use a parameter space described by observed $g-r$ colour and the time taken to rise from non-detection to peak $r$-band magnitude in the observer frame. Using rise times and colours from GP lightcurves, we populate this parameter space with the sample of 72 RETs from the \citetalias{Pursiainen2018} method, updated with 20 extra objects found using that method in the fifth season of DES. We add spectroscopically confirmed SNe of types Ia, Ibc, II, and SLSNe observed by DES in order to verify that they are rejected by the search method. We keep objects passing the following selection requirements (cuts):
\begin{enumerate}
\item The transient was detected in only one DES-SN observing season.
\item Maximum observed brightness in both $g$- and $r$-bands was brighter than 24 mag (in the eight `shallow' DES-SN fields) or 25 mag (in the two `deep' fields), as in \citetalias{Pursiainen2018}.
%\item $g$- and $r$-band observations used for the colour had to be taken within 2 days of each other in the observer frame.
\end{enumerate}
